\thetitle

### 2023-06-07 canonicitymodel.tex
\setc
Explain another notation for dependent function.
Make colorbox consistent.
Save. Fixing overflow.
Fix overflow.
Save.

  \denotesC{\goodCtx{\Gm}{k}} \text{ is a function }&: \setc{\gm}{\goodSub{\cdot}{\gm}{\Gm}} \to \Set{k} \quad \text{ (i.e., sets indexed by closed substitution)}\\

@@

  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \setc{ (\gamma' , t') }{ \gm' \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t' \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  } \\

@@

  & \text{Since $\denotesC{t}(\gm)(\gm')$ witnesses } \sub{t}{\gm} \equiv \eqrefl{\sub{u}{\gm}} \text{ and } \sub{u}{\gm} \equiv \sub{v}{\gm} \\

@@

These four sets are collecting the glued interpretation.
For each well-formed type $\goodType{\Gm}{T}{j}$, we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$;
for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$;
etc. The audience should reminisce these structure with the one appearing at the beginning of the MLTT canonicity model.

@@

  \colorbox{blue!10}{(\Gsub{\denotesCC{a}}{\Glued{\gm}})}~
  \colorbox{green!10}{(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})} \\

@@

\begin{align*}
  \text{we define } \RecSig{}{} &\text{ by induction on the signature, via } RS \\
  RS~"nil"~R &= \denotesCC{\LSigEmp} \\
  RS~((A, B) :: tl)~R &= \GLSigAdd{RS~tl~R}{\Glued{\pi_2}}{\GCaseSig{A}{B}{R}} \\ 
  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{"R"}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{"R"}} \\
  \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg 
\end{align*}
\begin{align*}
  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}}(\gm)(\gm') = \CWrec \quad &\Gsub{\denotesCC{\wsig}}{\Glued{\gm}} \\ 
  & (\lambda w. \denotesC{"R"}(\gm)(\gm')(\Wrec{\sub{\wsig}{\gm}}{\sub{\lkg}{\gm}}{w}))\\
  &f^r\\ &\sub{t}{\gm}\\ 
  &(\denotesC{t}(\gm)(\gm')) 
\end{align*}
\begin{align*}  
  \text{ where } \quad f^r~j~\Glued{a}~b~b^c &= 
  \texttt{let}~\Glued{\rho}  \in \GTm{
    \colorbox{blue!15}{(\Glued{\cdot} \Glued{,} \colorbox{blue!5}{\Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}}})}
    }{
      \colorbox{green!10}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})}
    } 
  \\
  &\texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^c)~\texttt{ in } \\
  & \Gmodel{
    \Gsub{
      \colorbox{blue!27}{\Gapp{
        \colorbox{blue!18}{\Gsub{
          \colorbox{blue!9}{\Gapp{
            \colorbox{blue!0}{\GRecproj{j}{\Gsub{\denotesCC{\lkg}}{\Glued{\gm}}}}
            }}
          }{
            \GSubstExt{\GSubstId}{\Glued{a}}
          }}}}
      }{
        \GSubstExt{\GSubstId}{
          \colorbox{green!10}{\Glam{\Glued{\rho}}}
          }
      }
    }(\epsilon)(\star) \\
  & \quad 
   \text{ given } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \\
   & \quad \text{ where }\Glued{R} = \denotesCC{"R"}, \Glued{B} = \Gwsigproj{j}{2}{\denotesCC{\wsig}}
\end{align*}

The semantic interpretations above are defined together with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$)
%
\begin{align*}
  "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
    \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\
    & \to \Glued{b} \in (\GTm{
      \colorbox{blue!15}{(\Glued{\cdot} \Glued{,} 
      \colorbox{blue!5}{\Gsub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}} )} 
      }{
        \colorbox{green!10}{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}
        }) \\
    & \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
\end{align*}
and its eliminator $\CWrec$
\begin{align*}
  \CWrec \  &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
    (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
    & \to 
    \Bigg( j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
    & \quad \quad \to 
    \setc{b}{\goodTerm{({\cdot} , 
    \colorbox{blue!10}{\sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}}
    ) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
    &  \quad \quad \to  
      \left( {\Glued{\gm} \in \GSub{\Glued{\cdot}}{
        \colorbox{blue!11}{(\Glued{\cdot} \Glued{,} 
    \colorbox{blue!1}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}}
    }\to P~(\sub{b}{ \gm }) \right) \\
    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) \Bigg) \\
    & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~\Glued{\wsig}~P~f~(\sub{b}{\gm})~(\Gmodel{\Glued{b}}~\Glued{\gm}))
\end{align*}

>>>>

  \denotesC{\goodCtx{\Gm}{k}} \text{ is a function }&: \setc{\gm}{\goodSub{\cdot}{\gm}{\Gm}} \to \Set{k} \quad \\
   & \text{ (i.e., sets indexed by closed substitution)}\\

@@

  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \setc{ (\gamma' , t') }{ \gm' \in \denotesC{{\Gm}}(\pi_1 \gm_t), t' \in \denotesC{{\sub{T}{\gm}}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  } \\

@@

   \text{Since $\denotesC{t}(\gm)(\gm')$ witnesses } &\sub{t}{\gm} \equiv \eqrefl{\sub{u}{\gm}} \text{ and } \sub{u}{\gm} \equiv \sub{v}{\gm} \\

@@

These four sets are collecting the glued interpretation.
For each well-formed type $\goodType{\Gm}{T}{j}$, we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$;
for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$;
etc. These structures should remind the reader of the ones appearing at the beginning of the MLTT canonicity model.

@@

  \colorbox{blue!9}{(\Gsub{\denotesCC{a}}{\Glued{\gm}})}~
  \colorbox{green!9}{(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})} \\

@@

\begin{align*}
  \text{we define } \RecSig{}{} &\text{ by induction on the signature, via } RS \\
  RS~"nil"~R &= \denotesCC{\LSigEmp} \\
  RS~((A, B) :: tl)~R &= \GLSigAdd{RS~tl~R}{\Glued{\pi_2}}{\GCaseSig{A}{B}{R}} \\ 
  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{"R"}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{"R"}} \\
  \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg 
\end{align*}
\begin{align*}
  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}}(\gm)(\gm') = \CWrec \quad &\Gsub{\denotesCC{\wsig}}{\Glued{\gm}} \\ 
  & (\lambda w. \denotesC{"R"}(\gm)(\gm')(\Wrec{\sub{\wsig}{\gm}}{\sub{\lkg}{\gm}}{w}))\\
  &f^r\\ &\sub{t}{\gm}\\ 
  &(\denotesC{t}(\gm)(\gm')) 
\end{align*}
\begin{align*}  
  \text{ where } \quad f^r~j~\Glued{a}~b~b^c &= 
  \texttt{let}~\Glued{\rho}  \in \GTm{
    \colorbox{blue!17}{(\Glued{\cdot} \Glued{,} \colorbox{blue!9}{\Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}}})}
    }{
      \colorbox{green!9}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})}
    } 
  \\
  &\texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^c)~\texttt{ in } \\
  & \Gmodel{
    \Gsub{
      \colorbox{blue!27}{\Gapp{
        \colorbox{blue!17}{\Gsub{
          \colorbox{blue!9}{\Gapp{
            \colorbox{blue!1}{\GRecproj{j}{\Gsub{\denotesCC{\lkg}}{\Glued{\gm}}}}
            }}
          }{
            \GSubstExt{\GSubstId}{\Glued{a}}
          }}}}
      }{
        \GSubstExt{\GSubstId}{
          \colorbox{green!9}{\Glam{\Glued{\rho}}}
          }
      }
    }(\epsilon)(\star) \\
  & \quad 
   \text{ given } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \\
   & \quad \text{ where }\Glued{R} = \denotesCC{"R"}, \Glued{B} = \Gwsigproj{j}{2}{\denotesCC{\wsig}}
\end{align*}

The semantic interpretations above are defined together with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$)

(Note: we also use $(a \in A \to B(a))$ as another notation for dependent function $\prod_{a \in A} B(a)$).
%
\begin{align*}
  "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \setc{t}{\goodTerm{\cdot}{t}{\El{\wcode{\wsig}}}} \to \Set{i + 1} \quad "where" \\
    \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\
    & \to \Glued{b} \in (\GTm{
      \colorbox{blue!17}{(\Glued{\cdot} \Glued{,} 
      \colorbox{blue!9}{\Gsub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}} )} 
      }{
        \colorbox{green!9}{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}
        }) \\
    & \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
\end{align*}
and its eliminator $\CWrec$
\begin{align*}
  \CWrec \  &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
    (P : \setc{t}{\goodTerm{\cdot}{t}{\El{\wcode{\wsig}}}} \to \Set{k})  \\
    & \to 
    \Bigg( j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
    & \quad \quad \to 
    \setc{b}{\goodTerm{({\cdot} , 
    \colorbox{blue!9}{\sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}}
    ) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
    &  \quad \quad \to  
      \left( {\Glued{\gm} \in \GSub{\Glued{\cdot}}{
        \colorbox{blue!17}{(\Glued{\cdot} \Glued{,} 
    \colorbox{blue!9}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}}
    }\to P~(\sub{b}{ \gm }) \right) \\
    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) \Bigg) \\
    & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~\Glued{\wsig}~P~f~(\sub{b}{\gm})~(\Gmodel{\Glued{b}}~\Glued{\gm}))
\end{align*}

### 2023-06-07 introduction.tex
Save.

Integrating family polymorphism into a dependent type theory for
logical reasoning, however, poses significant technical challenges.
As we analyze, pillars of dependent type theories—including
inductive types, definitional equality, and logical consistency—are
all inimical to the kind of extensibility and family polymorphism
found in existing OO language designs.
Thus, our contributions include novel design recipes for dealing with
these challenges and foundational meta\-theoretical guarantees on the
underlying logic.
%
Specifically, we make the following contributions.

>>>>

Integrating family polymorphism into a dependent type theory for
logical reasoning, however, poses significant technical challenges.
As our analysis
will show, pillars of dependent type theories—including
inductive types, definitional equality, and logical consistency—are
all inimical to the kind of extensibility and family polymorphism
found in existing OO language designs.
Thus, our contributions include novel design recipes for dealing with
these challenges and foundational meta\-theoretical guarantees on the
underlying logic.
%
Specifically, we make the following contributions.
